---
title: "DIID"
author: "KonstantinPopadin"
date: "8/19/2024"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1 Read data, derive DeficitOfFragility

```{r}
# Set the file path
file_path <- "2_Derived/Total_DIID_count.csv"
data <- read.table(file_path, sep = ';', header = TRUE)

# Display the first few rows of the data
names(data)

summary(data$DIIDFragilityScore)

data$DIID.DeficitOfFragility = data$D.I..IDcount - data$DI..I.Dcount # ALTER - FRAGILE: the higher the score the higher the deficit of fragile elements ( due to potential selection)
summary(data$DIID.DeficitOfFragility)
```

## 2 DIIDFragilityScore (==number of fragile / number of alternative structures)
### we expect that FragilityScore should decrease (due to selection) in case of: 
### (i) long repeats (so we expect negative correlations between FragilityScore and length);
### (ii) short intervals 1 and 3 (so we expect positive correlations between FragilityScore and interval length)
### in a set of univariate and multivariate analyses below we observe exactly this.

```{r}

summary(data$DirectRepeatLength)
summary(data$InvertedRepeatLength)

### UNIVARIATE WITH DIIDFragilityScore

cor.test(data$DIIDFragilityScore,data$DirectRepeatLength, method = 'spearman')   
plot(data$DIIDFragilityScore,data$DirectRepeatLength)

cor.test(data$DIIDFragilityScore,data$InvertedRepeatLength, method = 'spearman') 
plot(data$DIIDFragilityScore,data$InvertedRepeatLength)

cor.test(data$DIIDFragilityScore,data$LengthOfInterval1, method = 'spearman') 
plot(data$DIIDFragilityScore,data$LengthOfInterval1)

cor.test(data$DIIDFragilityScore,data$LengthOfInterval2, method = 'spearman')
plot(data$DIIDFragilityScore,data$LengthOfInterval2)

cor.test(data$DIIDFragilityScore,data$LengthOfInterval3, method = 'spearman')
plot(data$DIIDFragilityScore,data$LengthOfInterval3)

### MULTIVARIATE

summary(lm(data$DIIDFragilityScore ~ scale(data$DirectRepeatLength) + scale(data$InvertedRepeatLength) + scale(data$LengthOfInterval1) + scale(data$LengthOfInterval2) + scale(data$LengthOfInterval3)))

summary(lm(data$DIIDFragilityScore ~ scale(data$DirectRepeatLength) + scale(data$InvertedRepeatLength) + scale(data$LengthOfInterval1) + scale(data$LengthOfInterval3)))

```

## 3 DIID.DeficitOfFragility (== number of alternative structures - number of fragile structures)
### we expect that DeficitOfFragility (as a result of selection) should be more pronounced (stronger), when: 
### (i) repeats are long  (so we expect positive correlations between DeficitOfFragility and length);
### (ii) intervals 1 and 3 are short (so we expect negative correlations between DIID.DeficitOfFragility and interval length)
### in a set of univariate (not always) and in multivariate analyses below we observe exactly this.

```{r}
cor.test(data$DIID.DeficitOfFragility,data$DirectRepeatLength, method = 'spearman')   
plot(data$DIID.DeficitOfFragility,data$DirectRepeatLength)

cor.test(data$DIID.DeficitOfFragility,data$InvertedRepeatLength, method = 'spearman') 
plot(data$DIID.DeficitOfFragility,data$InvertedRepeatLength)

cor.test(data$DIID.DeficitOfFragility,data$LengthOfInterval1, method = 'spearman') 
plot(data$DIID.DeficitOfFragility,data$LengthOfInterval1)

cor.test(data$DIID.DeficitOfFragility,data$LengthOfInterval2, method = 'spearman')
plot(data$DIID.DeficitOfFragility,data$LengthOfInterval2)

cor.test(data$DIID.DeficitOfFragility,data$LengthOfInterval3, method = 'spearman')
plot(data$DIID.DeficitOfFragility,data$LengthOfInterval3)

### MULTIVARIATE

summary(lm(data$DIID.DeficitOfFragility ~ scale(data$DirectRepeatLength) + scale(data$InvertedRepeatLength) + scale(data$LengthOfInterval1) + scale(data$LengthOfInterval2) + scale(data$LengthOfInterval3)))

summary(lm(data$DIID.DeficitOfFragility ~ scale(data$DirectRepeatLength) + scale(data$InvertedRepeatLength) + scale(data$LengthOfInterval1) + scale(data$LengthOfInterval3)))

```

## 4: DG4IIDFragilityScore (==number of fragile with GQ / number of alternative structures)
### we expect that FragilityScore should decrease (due to selection) in case of: 
### (i) long repeats (so we expect negative correlations between FragilityScore and length);
### (ii) short intervals 1 and 3 (so we expect positive correlations between FragilityScore and interval length)
### in a set of univariate and multivariate analyses below we observe exactly this.

```{r}

summary(data$DirectRepeatLength)
summary(data$InvertedRepeatLength)
summary(data$DG4IIDFragilityScore)

### UNIVARIATE WITH DIIDFragilityScore

cor.test(data$DG4IIDFragilityScore,data$DirectRepeatLength, method = 'spearman')   
plot(data$DG4IIDFragilityScore,data$DirectRepeatLength)

cor.test(data$DG4IIDFragilityScore,data$InvertedRepeatLength, method = 'spearman') 
plot(data$DG4IIDFragilityScore,data$InvertedRepeatLength)

cor.test(data$DG4IIDFragilityScore,data$LengthOfInterval1, method = 'spearman') 
plot(data$DG4IIDFragilityScore,data$LengthOfInterval1)

cor.test(data$DG4IIDFragilityScore,data$LengthOfInterval2, method = 'spearman')
plot(data$DG4IIDFragilityScore,data$LengthOfInterval2)

cor.test(data$DG4IIDFragilityScore,data$LengthOfInterval3, method = 'spearman')
plot(data$DG4IIDFragilityScore,data$LengthOfInterval3)

### MULTIVARIATE

summary(lm(data$DG4IIDFragilityScore ~ scale(data$DirectRepeatLength) + scale(data$InvertedRepeatLength) + scale(data$LengthOfInterval1) + scale(data$LengthOfInterval2) + scale(data$LengthOfInterval3)))

summary(lm(data$DG4IIDFragilityScore ~ scale(data$DirectRepeatLength) + scale(data$InvertedRepeatLength) + scale(data$LengthOfInterval1) + scale(data$LengthOfInterval3)))

```



Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
